Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-04-20 draft_p2.tex
Divide AMS subsection into two subsubsections

% Example: OU process
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling. The orange line represents the evolution of the maximum over resampled trajectories as a function of the computational cost $C_{TAMS}$ (iterations of the the selection-mutation step). The solid blue line is the analytical solution for the return time of amplitude $a$ for the \ac{ou} process~\cite{lestang_computing_2018}.}
  \label{fig:comparaison_temps_de_retour}
\end{figure}
We first describe results of the \ac{tams} on a simple case.
We temporarily let fluid dynamics aside and consider a one-dimensional \ac{ou} process
\begin{equation}
  \label{eq:ou}
  \dot{x} = -x + \sqrt{2\epsilon}\eta (t)  
\end{equation}
We apply the \ac{tams} to simulate trajectories $\{x(t)\}_{0\leq t \leq T_a}$ with $\underset{0\leq t \leq T_a}{\max} x(t) \geq a$ with $a$ very large with respect to typical fluctuations of $x(t)$.
For the sake of simplicity, a single trajectory is resampled at each iteration of the \ac{tams}.
We denote by $a_j$ the maximum of the resampled trajectory at iteration $j$.
The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
Figure~\ref{fig:compraison_temps_de_retour} compares the typical computational cost required to sample fluctuations above a given amplitude $a$ using the \ac{tams} and a direct sampling approach. In the latter, the typical computational cost is simply the return time $r(a)$.
illustrates that the successice resamplings of the \ac{tams} algorithm leads to trajectories displaying extreme fluctuations $x \geq a$.
More importantly, these very rare trajectories are sampled for a computational cost that several order of magnitude lower than the return time of the corresponding fluctuations.

@@

% Application to turbulent flow past an obstacle
Can the \ac{tams} achieve similar results when applied to complex chaotic dynamics, such as turbulent flows around objects ?
To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
Our aim is to sample trajectories that display extreme fluctuations of the drag $f_d$ acting on the square obstacle.
However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
As a consequence the choice of the score function $\xi (\mathbf{x}(t),t)$ is not so clear.
In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
For such complex dynamics however, there is no reasons for the drag itself to be a good measure of how far is a given state to regions of phase space where the drag is extreme.
Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, which is precisely what we are trying to achieve.

>>>>

\subsubsection{Illustration of the \ac{ams} on a simple case: the \acl{ou} process}
\label{sec:ams_ou}
% Example: OU process
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{AMS_OU/AMS_OU.png}
    \caption{Illustration of the efficiency of the \ac{tams} algorithm with respect to direct sampling. The orange line represents the evolution of the maximum over resampled trajectories as a function of the computational cost $C_{TAMS}$ (iterations of the the selection-mutation step). The solid blue line is the analytical solution for the return time of amplitude $a$ for the \ac{ou} process~\cite{lestang_computing_2018}.}
  \label{fig:comparaison_temps_de_retour}
\end{figure}
We temporarily let fluid dynamics aside and consider a one-dimensional \ac{ou} process
\begin{equation}
  \label{eq:ou}
  \dot{x} = -x + \sqrt{2\epsilon}\eta (t)  
\end{equation}
We apply the \ac{tams} to simulate trajectories $\{x(t)\}_{0\leq t \leq T_a}$ with $\underset{0\leq t \leq T_a}{\max} x(t) \geq a$ with $a$ very large with respect to typical fluctuations of $x(t)$.
For the sake of simplicity, a single trajectory is resampled at each iteration of the \ac{tams}.
We denote by $a_j$ the maximum of the resampled trajectory at iteration $j$.
The computational cost of the algorithm after iteration $j$ is the computational cost of simulating the $N$ initial trajectories, as well as the resampling of the $j$ trajectories.
Figure~\ref{fig:compraison_temps_de_retour} compares the typical computational cost required to sample fluctuations above a given amplitude $a$ using the \ac{tams} and a direct sampling approach. In the latter, the typical computational cost is simply the return time $r(a)$.
illustrates that the successice resamplings of the \ac{tams} algorithm leads to trajectories displaying extreme fluctuations $x \geq a$.
More importantly, these very rare trajectories are sampled for a computational cost that several order of magnitude lower than the return time of the corresponding fluctuations.

@@

\subsubsection{The \ac{ams} for extreme drag fluctuations}`
\label{sec:ams_drag}
% Application to turbulent flow past an obstacle
Can the \ac{tams} achieve similar results when applied to complex chaotic dynamics, such as turbulent flows around objects ?
To answer this question, we consider again the two-dimensional test flow introduced in section~\ref{sec:test_flow}.
Our aim is to sample trajectories that display extreme fluctuations of the drag $f_d$ acting on the square obstacle.
However, by contrast with a simple \ac{ou} process, the phase space is higly-dimensional and the dynamics complex.
As a consequence the choice of the score function $\xi (\mathbf{x}(t),t)$ is not so clear.
In the following we use the \ac{tams} with $\xi \equiv f_d$, which is the most simple choice for the score function.
For such complex dynamics however, there is no reasons for the drag itself to be a good measure of how far is a given state to regions of phase space where the drag is extreme.
Unfortunately, the choice of better score functions rely on a deep understanding of the physics of the rare events of interest, which is precisely what we are trying to achieve.
